---
title: "STAT 481 - Project 1"
author: "Thomas Gierach"
date: "3/30/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The experiment is to develop one's own rating system and find which cereal is the most healthy for you.Try to predict the cereal rating based on the nutrition facts and other observable characteristics. What variables best predict the rating? Can you quantify how good your predictive model is?

rating =  protein + sodium + fiber  + sugars + vitamins + weight + cups.



```{r, echo = TRUE}
cereals = read.csv("CerealsRating.csv")
str(cereals)
summary(cereals)
cereals
cereals2 <- cereals[,2:length(cereals)]
cereals2

```
```{r, echo = TRUE}

## Normality Check
attach(cereals2)
MLR <- lm(rating~protein+sodium+fiber+sugars+vitamins+weight+cups)
summary(MLR)
qqnorm(MLR$residual)
qqline(MLR$residual)
shapiro.test(MLR$residual)

## Equal Variance Check
plot(MLR)

#The points on the residuals vs fitted are skewed to the right

## Linearity Check
	par(mfrow=c(3,3))
	plot(x = protein, y = MLR$residual, col = "red")
	abline(h=0)
	plot(x = sodium, y = MLR$residual, col = "blue")
	abline(h=0)
	plot(x = fiber, y = MLR$residual, col = "gray")
	abline(h=0)
	plot(x = sugars, y = MLR$residual, col = "green")
	abline(h=0)
	plot(x = vitamins, y = MLR$residual, col = "orange")
	abline(h=0)
	plot(x = weight, y = MLR$residual, col = "yellow2")
	abline(h=0)
	plot(x = cups, y = MLR$residual, col = "violet")
	abline(h=0)
	
	#protein has very little data at 5
	#sodium is skewed to the left
	#fiber is skewed to the right
	#sugars is random
	#vitamins have a big gap between 25 and 100
	#weight is mostly at 1.0
	#cups has no pattern
	
	## Independence Check (Durbin-Watson): Not required since not time-series data
	dwtest(MLR, alternative = "two.sided")
```
```{r echo = TRUE}
## Transformation on Y
library(MASS)
trans <- boxcox(rating~protein+sodium+fiber+sugars+vitamins+weight+cups,
	data=cereals2,
	plotit=F, #get boxcox instead of the plot
	lambda = seq(-3, 3, by=0.125))
maxyentry <- which.max(trans$y) # obtain index number of largest y
trans$x[maxyentry] # obtain lambda value

## Add new column to dataframe with transformed variable
cereals2$logY <- log(cereals2$rating)
cereals2$logprotein <- log(cereals2$protein)
cereals2$logsodium <- log(cereals2$sodium)
cereals2$logfiber <- log(cereals2$fiber)
cereals2$logvitamins <- log(cereals2$vitamins)
cereals2$logweight <- log(cereals2$weight)
```

``` {r echo = TRUE}
## Full Model with Y-Transformation
library(car)
attach(cereals2)
MLR2 <- lm(logY~protein+sodium+fiber+sugars+vitamins+weight+cups)
summary(MLR2) # gives partial t-test values and regression F value
anova(MLR2) # provides ANOVA table for overall model

## Normality Check
par(mfrow=c(1,1))
qqnorm(MLR2$residual)
qqline(MLR2$residual)
shapiro.test(MLR2$residual)

## Equal Variance Check (top left corner)
par(mfrow=c(2,2))
plot(MLR2)

## Linearity Check
par(mfrow=c(3,3))
plot(x = protein, y = MLR2$residual, col = "red")
abline(h=0)
plot(x = sodium, y = MLR2$residual, col = "blue")
abline(h=0)
plot(x = fiber, y = MLR2$residual, col = "gray")
abline(h=0)
plot(x = sugars, y = MLR2$residual, col = "green")
abline(h=0)
plot(x = vitamins, y = MLR2$residual, col = "orange")
abline(h=0)
plot(x = weight, y = MLR2$residual, col = "yellow2")
abline(h=0)
plot(x = cups, y = MLR2$residual, col = "violet")
abline(h=0)

## VIF Check
vif(MLR2)
	
MLR3 <- lm(rating~protein+sodium+fiber+sugars+vitamins+cups)
summary(MLR3)
MLR4 <- lm(rating~protein+sodium+fiber+sugars+cups)
summary(MLR4)
MLR5 <- lm(rating~protein+sodium+fiber+sugars)
summary(MLR5)
anova(MLR5)
MLR6 <- lm(rating~sodium+fiber+sugars)
summary(MLR6)
anova(MLR6)

#Normality Check
qqnorm(MLR6$residual)
qqline(MLR6$residual)
shapiro.test(MLR6$residual)

## Equal Variance Check
plot(MLR6)

## Linearity Check
	par(mfrow=c(1,3))

	plot(x = cereals2$sodium, y = MLR6$residual, col = "blue")
	abline(h=0)
	plot(x = cereals2$logfiber, y = MLR6$residual, col = "gray")
	abline(h=0)
	plot(x = sugars, y = MLR6$residual, col = "green")
	abline(h=0)
	
	## Independence Check (Durbin-Watson): Not required since not time-series data
	dwtest(MLR, alternative = "two.sided")
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
